---
title: "NFL Scrape Test Icing Kickers"
author: "Sam Silva"
date: "11/20/2016"
output:
  html_document: default
  word_document: default
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  cache = TRUE,
  tidy = TRUE
)
require(pacman)
```

```{r loadpacs, echo=TRUE}
# Must install the devtools package using the below commented out code
# install.packages("devtools")
  p_load(devtools)
  #devtools::install_github(repo = "maksimhorowitz/nflscrapR")
# Load the package
  p_load(nflscrapR)
```


Get the game information for all NFL games from 2009 to 2015.  NFL.com doesn't want to serve 2007 and 2008.

Change the `` to ``` at the beginning and end of the chunk to make it executable.  The chunk can be made non-executable once the dataframe has been created and saved for reloading.

```{r grabgames, echo=TRUE}
  # NFL.COM seems not to want to serve years 2007 and 2008
  years <- c(2009:2015)
  for (i in 1:length(years)){
    print(years[i])
    assign(paste0("gameids", years[i]), extracting_gameids(years[i], playoffs = FALSE))
  }
```

Now grab all the play information for all of the games that were found above.

Change the `` to ``` at the beginning and end of the chunk to make it executable. The chunk can be made non-executable once the dataframe has been created and saved for reloading. 

```{r, getplays, echo=TRUE}
  years <- c(2009:2015)
  rm(plays)
  print(years[1])
  gameids <- eval(as.symbol(paste0("gameids", years[1])))
  plays <- game_play_by_play(gameids[1])
  for(i in 2:length(gameids)){
    tempplays <- game_play_by_play(gameids[i])
    plays <- rbind(plays, tempplays)
    rm(tempplays)
    print(c(years[1], paste(i/length(gameids)*100,"%")))
  }
  for(j in 2:length(years)){
    print(years[j])
    gameids <- eval(as.symbol(paste0("gameids", years[j])))  
    for(i in 1:length(gameids)){
      tempplays <- game_play_by_play(gameids[i])
      plays <- rbind(plays, tempplays)
      rm(tempplays)
      print(c(years[j], paste(i/length(gameids)*100,"%")))
    }
  }
  dim(plays)
  head(plays)
  write.csv(plays,"plays.csv")
```

We can create a new dataframe that contains "Timeout" and "Field Goal" events.

Change the `` to ``` at the beginning and end of the chunk to make it executable.  The chunk can be made non-executable once the dataframe has been created and saved for reloading.

```{r subsetFGandTO, echo=TRUE}
  table(plays$PlayType)
  FGsTOs <- subset(plays, PlayType %in% c("Timeout", "Field Goal"))
  # FGsTOs <- plays[plays$PlayType %in% c("Timeout", "Field Goal"),]
  table(FGsTOs$PlayType) 
  write.csv(FGsTOs,"FGsTOs.csv")
```

Determine if the kicker was iced and subset the data to get only field goals (FGs).

```{r subsetFG, echo=TRUE}
  p_load(dplyr)
  p_load(stringr)

  ### Comment out the next line if you are building the plays above
  ### Change the file location to reflect where it is stored
  #load("D:/Fall16/FYS/NFLScrape/PlayByPlay/FYSNFLproject20161120.RData")
  #load("/Users/samuelsilva/Desktop/FYSNFLproject20161120.Rdata")
  
  ### Using plays
  temp = plays
  temp$desc1 = toupper(lag(temp$desc,1))
  temp$desc2 = toupper(lag(temp$desc,2))
  temp$time1 = lag(temp$time,1)
  temp$time2 = lag(temp$time,2)
  temp$PlayType1 = lag(temp$PlayType,1)
  temp$PlayType2 = lag(temp$PlayType,2)
  temp$DefensiveTeam = toupper(temp$DefensiveTeam)
  temp$Iced1 = rep(NA,nrow(temp))
  temp$Iced1 = (temp$PlayType1=="Timeout" & temp$PlayType=="Field Goal") & (temp$time1==temp$time) & !is.na(str_locate(temp$desc1,temp$DefensiveTeam)[,1])
  table(temp$Iced1)
   temp$Iced2 = (temp$PlayType2=="Timeout" & temp$PlayType=="Field Goal") & (temp$time2==temp$time) & !is.na(str_locate(temp$desc2,temp$DefensiveTeam)[,2])
   temp$Iced = temp$Iced1 | temp$Iced2
  xtabs(~Iced2, data=temp)
  xtabs(~Iced1+Iced2, data=temp)
  xtabs(~FieldGoalResult+(Iced1 | Iced2), data=temp)
  temp$obsNo = 1:nrow(temp)
  
  FGs = subset(temp, temp$PlayType=="Field Goal")
  FGs$FieldGoalDistance = as.numeric(FGs$FieldGoalDistance)
  FGs$FieldGoalGood = factor(FGs$FieldGoalResult=="Good")
  FGs$Iced = factor(FGs$Iced)
  FGs$Iced1 = factor(FGs$Iced1)
  FGs$Iced2 = factor(FGs$Iced2)
  xtabs(~FieldGoalGood+Iced, data=FGs)
 
  # Recode TRUE/FALSE as No/Yes
  FGs$Iced=factor(FGs$Iced, levels = c(FALSE,TRUE), labels = c("No","Yes"))
  FGs$FieldGoalGood=factor(FGs$FieldGoalGood, levels = c(FALSE,TRUE), labels = c("No","Yes"))
  
  #write.csv(FGs,"d:/fall16/fys/nflscrape/playbyplay/FGs.csv")
 
  # Remove temp since it is a bigger version of plays with subseting variables
  rm(temp)
  save.image("FYSNFLproject20161216.RData")
```

Add kicker

```{r addkicker, echo=TRUE}
 p_load(stringr)
 #tmpx = FGs[c(1,7,15),"desc"]
 getkicker <- function(x=tmpx){
    n <- length(x)
    kicker <- rep(NA,n)
    #print(paste("kicker=",kicker))
    for (i in 1:n){
      #print(paste("i=",i," of ",n))
      tmp <- strsplit(x[i], split="formation) ")
      tmpn <- length(tmp[[1]])
      #print(paste("tmp=", tmp))
      #print(paste("tmpn=",tmpn))
      if (tmpn==1){
        tmpk <- strsplit(tmp[[1]][1], split=") ")
        #print(paste("tmpk1=",tmpk))
        tmpk <- strsplit(tmpk[[1]][2],split=" ")
        #print(paste("tmpk2=",tmpk))
        kicker[i] <- tmpk[[1]][1]
        #print(paste("kicker=",kicker))
      }else{
        tmp <- strsplit(tmp[[1]][2], split=" ")
        #print(paste("tmp=",tmp))
        kicker[i] <- tmp[[1]][1]
        #print(paste("kicker=",kicker))
      } # else
    } # for
    return(str_trim(chartr(",.","  ",kicker),"both"))
 } #getkicker
 #getkicker(tmpx)
 FGs$kicker = getkicker(FGs$desc)
 FGs$A.Vinatieri = factor(FGs$kicker=="A Vinatieri",levels = c(FALSE,TRUE),labels = c("No","Yes"))
  FGs$D.Bailey = factor(FGs$kicker=="D Bailey",levels = c(FALSE,TRUE),labels = c("No","Yes"))
  FGs$M.Crosby = factor(FGs$kicker=="M Crosby",levels = c(FALSE,TRUE),labels = c("No","Yes"))
  xtabs(~kicker+M.Crosby, data=FGs)
```

Load packages needed to fit the logist regression models (lrm) and to generate predicted probability plots and nomograms.

```{r loadpacks, echo=TRUE}
  # Load a few packages
  p_load(Hmisc)
  p_load(xtable)
  p_load(lattice)
  p_load(rms)  ### Modern R replacement for the Design package
```

## Fitting Logistic Models

### Distance

The method of fitting models by finding the observed log-odds that is demonstrated above is tedious.  A model to test for the difference in odds of **FieldGoalGood** = "Y" as determined by **FieldGoalDistance** may be fitted using the **lrm** function.

```{r lrmdistance, echo=TRUE}
  # Define the data.frame removing variables that won't be used.
  FG = FGs[,c(1:19, 49:50, 62:65, 67:ncol(FGs))]
  dd = datadist(FG)
  options(datadist="dd")
  if ("FG" %in% search()) detach(FG)
  attach(FG)
  # Fit a logistic model of Hit as a function of Distance
  FG.lrm.dist=lrm(FieldGoalGood~FieldGoalDistance, x=TRUE, y=TRUE)
  FG.lrm.dist
  # Get a little information about the quality of the model
  anova(FG.lrm.dist)
```

Since the coefficient of **FieldGoalDistance** is negative, we see that the (log) odds of making a shot (**FieldGoalGood** = "Y") decreases as **FieldGoalDistance** increases.  This can most easily be seen using the odds ratios.  Comparisons against the minimum distance show the effect of moving farther back.  E.g. a FG from 60 yards is about 0.015 (0.010, 0.020) as likely to be made as one taken from 18 yrds.

```{r oddsdist, echo=TRUE}
  summary(FG.lrm.dist)
  summary(FieldGoalDistance, na.rm=TRUE)
  min(FieldGoalDistance, na.rm=TRUE)
  summary(FG.lrm.dist, FieldGoalDistance=c(min(FieldGoalDistance, na.rm=TRUE),30))
  summary(FG.lrm.dist, FieldGoalDistance=c(min(FieldGoalDistance, na.rm=TRUE),40))
  summary(FG.lrm.dist, FieldGoalDistance=c(min(FieldGoalDistance, na.rm=TRUE),50))
  summary(FG.lrm.dist, FieldGoalDistance=c(min(FieldGoalDistance, na.rm=TRUE),60))

```

The probability of making the FG at different distances may be plotted.

```{r plotdist, echo=TRUE}
  print(plot(Predict(FG.lrm.dist, fun=plogis), ylab="Probability of Making a Kick", xlab="Yards"))
```

We see that as the kick is taken from farther back, the probability of their making the FG decreases.  A nomogram may be helpful in estimating these probabilities.

```{r nomogramdist, echo=TRUE}
  nom = nomogram(FG.lrm.dist, fun=plogis)
  print(plot(nom))
```

### Icing

A model to test for the difference in odds of **FieldGoalGood** = "Y" as determined by **Iced** may be fitted using the **lrm** function.

```{r lrmhand, echo=TRUE}
  xtabs(~Iced, data=FGs)
  FG = FGs[,c(1:19, 49:50, 62:65, 67:ncol(FGs))]
  dd = datadist(FG)
  options(datadist="dd")
  if ("FG" %in% search()) detach(FG)
  attach(FG)
  FG.lrm.Iced=lrm(FieldGoalGood~Iced, x=TRUE, y=TRUE)
  FG.lrm.Iced
  anova(FG.lrm.Iced)
  summary(FG.lrm.Iced, Iced=c("No"))
  summary(FG.lrm.Iced, Iced=c("Yes"))
  plot(Predict(FG.lrm.Iced, fun=plogis), ylab="Probability of Making a Kick")
  nom = nomogram(FG.lrm.Iced, fun=plogis)
  plot(nom)
```

### Kicker

```{r lrmkicker, echo=TRUE}

 ### Code to the next comment subsets the data for kickers with ge 20 attempts
 ### who were iced more than 4 times
 kickertable = table(FGs$kicker)
 kickertable
 FGs$kickerge20 = FGs$kicker %in% names(kickertable[kickertable >= 20])
 kickericedtable = table(FGs$kicker, FGs$Iced)
 kickericedtable
 FGs$kickericed = FGs$kicker %in% dimnames(kickericedtable[kickericedtable[,2] > 4,])[[1]]
 xtabs(~kickericed+Iced, data=FGs)
 FG = subset(FGs, FGs$kickericed)
 xtabs(~kicker+Iced, data=FG)
 ### End of subsetting
 
 # Refit distance to see if things changed
  # Define the data.frame using subsetted data.frame removing variables that     
  # won't be used.
  FG = FG[,c(1:19, 49:50, 62:65, 67:ncol(FG))]
  dd = datadist(FG)
  options(datadist="dd")
  while("FG" %in% search()) { detach(FG) }
  attach(FG)
  
  # Fit a logistic model of Hit as a function of Distance
  FG.lrm.dist=lrm(FieldGoalGood~FieldGoalDistance, x=TRUE, y=TRUE)
  FG.lrm.dist
  # Get a little information about the quality of the model
  anova(FG.lrm.dist)
  # Odds ratios
  summary(FG.lrm.dist)
  summary(FieldGoalDistance, na.rm=TRUE)
  min(FieldGoalDistance, na.rm=TRUE)
  summary(FG.lrm.dist, FieldGoalDistance=c(min(FieldGoalDistance, na.rm=TRUE),30))
  summary(FG.lrm.dist, FieldGoalDistance=c(min(FieldGoalDistance, na.rm=TRUE),40))
  summary(FG.lrm.dist, FieldGoalDistance=c(min(FieldGoalDistance, na.rm=TRUE),50))
  summary(FG.lrm.dist, FieldGoalDistance=c(min(FieldGoalDistance, na.rm=TRUE),60))
    print(plot(Predict(FG.lrm.dist, fun=plogis), ylab="Probability of Making a Kick", xlab="Yards"))
  # Nomogram
  nom = nomogram(FG.lrm.dist, fun=plogis)
  print(plot(nom))
  
  # Refit Iced model
  FG.lrm.Iced=lrm(FieldGoalGood~Iced, x=TRUE, y=TRUE)
  FG.lrm.Iced
  anova(FG.lrm.Iced)
  summary(FG.lrm.Iced, Iced=c("No"))
  summary(FG.lrm.Iced, Iced=c("Yes"))
  plot(Predict(FG.lrm.Iced, fun=plogis), ylab="Probability of Making a Kick")
  nom = nomogram(FG.lrm.Iced, fun=plogis)
  plot(nom)
  
  # Fit kicker model
  FG.lrm.kicker=lrm(FieldGoalGood~kicker, x=TRUE, y=TRUE)
  FG.lrm.kicker
  anova(FG.lrm.kicker)
  summary(FG.lrm.kicker, kicker=c("A Vinatieri"))
  summary(FG.lrm.kicker, kicker=c("M Crosby"))
  plot(Predict(FG.lrm.kicker, fun=plogis), ylab="Probability of Making a Kick")
  nom = nomogram(FG.lrm.kicker, fun=plogis)
  plot(nom)
  
  # Fit kicker*Iced model
  FG.lrm.kickerIced=lrm(FieldGoalGood~kicker*Iced, x=TRUE, y=TRUE)
  FG.lrm.kickerIced
  anova(FG.lrm.kickerIced)
  summary(FG.lrm.kickerIced, kicker=c("A Vinatieri"), Iced=c("No"))
  plot(Predict(FG.lrm.kickerIced, fun=plogis, kicker=c("A Vinatieri","M Crosby"), Iced=c("No","Yes")), ylab="Probability of Making a Kick")
  nom = nomogram(FG.lrm.kickerIced, fun=plogis)
  plot(nom)
  
  # Fit kicker*FieldGoalDistance model
  FG.lrm.kickerDist=lrm(FieldGoalGood~kicker*FieldGoalDistance, x=TRUE, y=TRUE)
  FG.lrm.kickerDist
  anova(FG.lrm.kickerDist)
  summary(FG.lrm.kickerDist, kicker=c("A Vinatieri"))
  plot(Predict(FG.lrm.kickerDist, fun=plogis, kicker=c("A Vinatieri","M Crosby")), ylab="Probability of Making a Kick")
  nom = nomogram(FG.lrm.kickerDist, fun=plogis)
  plot(nom)
  
  # Fit Iced*FieldGoalDistance model
  FG.lrm.IcedDist=lrm(FieldGoalGood~Iced*FieldGoalDistance, x=TRUE, y=TRUE)
  FG.lrm.IcedDist
  anova(FG.lrm.IcedDist)
  summary(FG.lrm.IcedDist, Iced=c("No"))
  plot(Predict(FG.lrm.IcedDist, fun=plogis), ylab="Probability of Making a Kick")
  nom = nomogram(FG.lrm.IcedDist, fun=plogis)
  plot(nom)
```

